---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# USFSAevents

<!-- badges: start -->
<!-- badges: end -->

The goal of USFSAevents is to scrape figure skating event data from the US Figure Skating Association (USFSA) website, perform summary statistics, and create relevant visual aids in comparing various competitions.

## Installation

You can install the released version of USFSAevents from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("USFSAevents")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ayak9/fs_repo")
```
## Example

This is a basic example:

```{r example}
library(USFSAevents)
Event1 = anycomp('https://ijs.usfigureskating.org/leaderboard/results/2018/26192/SEGM036.html')
```

